Final Ratings for Genome Biology

### 2018-10-25 datasets/microbiome/microbiome-enigma.Rmd

Here we analyze a subset of the ENIGMA data ([Smith, M. and Rocha, A. et al 2015](http://dx.doi.org/10.1128/mBio.00326-15)).
The dataset includes water samples from 67 wells with well-characterized biochemical metadata (i.e. pH, heavy metal concentrations, etc).
We acquired the data from a collaborator and have put it in the `benchmark-fdr/datasets/microbiome/DATA/enigma` folder.

@@

```{r, enigma-download-data}
if (!file.exists(file.path(enigma_dir, "enigma.metadata.txt"))) {
    download.file("https://www.dropbox.com/s/0r1zjp26u10l0p4/enigma.metadata.txt?dl=1",
                  destfile = file.path(enigma_dir, "enigma.metadata.txt"))
}

if (!file.exists(file.path(enigma_dir, "enigma.otu_table_resampled_updated_r.txt.gz"))) {
    download.file("https://www.dropbox.com/s/u9qh86cd7c0vbn9/enigma.otu_table_resampled_updated_r.txt.gz?dl=1",
                  destfile = file.path(enigma_dir, "enigma.otu_table_resampled_updated_r.txt.gz"))
}
```

```{r, enigma-readdata}
## load OTU table and metadata
print(file.path(enigma_dir, "enigma.otu_table_resampled_updated_r.txt.gz"))
otu <- read.table(file.path(enigma_dir, "enigma.otu_table_resampled_updated_r.txt.gz"), 
                  sep='\t', header = TRUE, row.names = 1)
meta <- read.csv(file.path(enigma_dir, "enigma.metadata.txt"), sep='\t')

>>>>

Here we analyze a subset of the ENIGMA data ([Smith, M. and Rocha, A. et al 2015](https://doi.org/10.1128/mBio.00326-15)).
The dataset includes water samples from 67 wells with well-characterized biochemical metadata (i.e. pH, heavy metal concentrations, etc).
We'll download the processed OTU table and associated metadata from Zenodo and put it in the `data/enigma_results` folder.

@@

```{r, enigma-download-data}
if (!file.exists(file.path(enigma_dir, "enigma.metadata.txt"))) {
    download.file("https://zenodo.org/record/1455793/files/enigma.metadata.txt",
                  destfile = file.path(enigma_dir, "enigma.metadata.txt"))
}

if (!file.exists(file.path(enigma_dir, "enigma.otu_table_resampled_updated_r.txt.gz"))) {
    download.file("https://zenodo.org/record/1455793/files/enigma.otu_table_resampled_updated_r.txt.gz",
                  destfile = file.path(enigma_dir, "enigma.otu_table_resampled_updated_r.txt.gz"))
}
```

```{r, enigma-readdata}
## load OTU table and metadata
otu <- read.table(file.path(enigma_dir, "enigma.otu_table_resampled_updated_r.txt.gz"), 
                  sep='\t', header = TRUE, row.names = 1)
meta <- read.csv(file.path(enigma_dir, "enigma.metadata.txt"), sep='\t')

### 2018-10-25 datasets/microbiome/microbiome-goodrich.Rmd

Here we download and analyze the Goodrich obesity (OB) dataset
[(Goodrich et al., 2014)](http://dx.doi.org/10.1016/j.cell.2014.09.053).
The dataset includes fecal samples from 428 individuals with lean body mass index (BMI < 25),
and 185 obese individuals (BMI > 30). We'll download the processed OTU tables from
Zenodo and unzip them in the `benchmark-fdr/datasets/microbiome/data` folder.

We expect few or no truly differentially abundant OTUs ([Finucane et al., 2014](http://dx.doi.org/10.1371/journal.pone.0084689) and [Walters et al., 2014](https://dx.doi.org/10.1016%2Fj.febslet.2014.09.039)), so this study provides a real-life "null" example.

>>>>

Here we download and analyze the Goodrich obesity (OB) dataset
[(Goodrich et al., 2014)](https://doi.org/10.1016/j.cell.2014.09.053).
The dataset includes fecal samples from 428 individuals with lean body mass index (BMI < 25),
and 185 obese individuals (BMI > 30). We'll download the processed OTU tables from
Zenodo and unzip them in the `data/ob_goodrich_results` folder.

We expect few or no truly differentially abundant OTUs ([Finucane et al., 2014](https://doi.org/10.1371/journal.pone.0084689) and [Walters et al., 2014](https://doi.org/10.1016%2Fj.febslet.2014.09.039)), so this study provides a real-life "null" example.

### 2018-10-25 datasets/microbiome/microbiome-schubert.Rmd

Here we download and analyze the Schubert diarrhea dataset [(Schubert et al., 2014)](http://dx.doi.org/10.1128/mBio.01021-14).
The dataset includes 154 healthy nondiarrheal, 93 Clostridium difficile infection (CDI) associated diarrheal,
and 89 non-CDI associated diarrheal stool samples. We'll download the processed OTU tables from
Zenodo and unzip them in the `benchmark-fdr/datasets/microbiome/data` folder.

@@

cdi_data <- file.path(datdir, "cdi_schubert_results.tar.gz")
cdi_dir <- file.path(datdir, "cdi_schubert_results")

>>>>

Here we download and analyze the Schubert diarrhea dataset [(Schubert et al., 2014)](https://doi.org/10.1128/mBio.01021-14).
The dataset includes 154 healthy nondiarrheal, 93 Clostridium difficile infection (CDI) associated diarrheal,
and 89 non-CDI associated diarrheal stool samples. We'll download the processed OTU tables from
Zenodo and unzip them in the `data/cdi_schubert_results` folder.

@@

cdi_data <- file.path(datdir, "cdi_schubert_results.tar.gz")
cdi_dir <- file.path(datdir, "cdi_schubert_results")
dir.create(cdi_dir, showWarnings = FALSE)

### 2018-10-25 manuscript_figures/manuscript_ratings.Rmd

```{r, fig.width = 6, fig.height = 4}
bind_rows(yeast = fdrcontrol_yeast, sims = fdrcontrol_sims, .id = "analysis") %>%
    ggplot(aes(x = analysis, y = pExceeds, color = Method, group = Method)) +
    geom_line(aes(linetype = Method), alpha = 1/2, position = position_dodge(width = 1/4)) +
    geom_point(position = position_dodge(width = 1/4)) + 
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_discrete(NULL) + 
    scale_y_continuous("Percent Settings Exceeding nominal FDR", labels = scales::percent) + 
    expand_limits(y = 0) +
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14)) +
    ggtitle("FDR control across simulations and yeast experiments") 
```

@@

```{r, fig.width = 6, fig.height = 4}
bind_rows(yeast = power_yeast, sims = power_sims, .id = "analysis") %>%
    dplyr::mutate(Method = reorder(Method, meanRank)) %>%
    ggplot(aes(x = Method, y = meanRank, color = Method, group = analysis)) +
    geom_line(aes(linetype = analysis), color = 'black', alpha = 1/4) +
    geom_point() + 
    scale_color_manual(values = col, guide = FALSE) +
    scale_x_discrete(NULL) + 
    scale_y_continuous("Mean TPR rank", breaks = 0:10) + 
    expand_limits(y = 1) +
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14)) +
    ggtitle("TPR rank across simulations and yeast experiments") 
```

@@

```{r, fig.width = 6, fig.height = 8}
bind_rows(yeast = robustness_yeast, sims = robustness_sims,
          casestudies = robustness_data, .id = "analysis") %>%
    dplyr::mutate(Method = reorder(Method, sdGain)) %>%
    ggplot(aes(x = Method, y = sdGain, color = Method, group = analysis)) +
    geom_line(aes(linetype = analysis), color = 'black', alpha = 1/4) +
    geom_point() + 
    scale_color_manual(values = col, guide = FALSE) +
    scale_x_discrete(NULL) + 
    scale_y_continuous("SD of percent relative to BH/qvalue") + 
    expand_limits(y = 1) +
    facet_grid(key ~ ., scales = "free") + 
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14)) +
    ggtitle("Variation across simulations, yeast experiments,\ncase studies")

@@

>>>>

```{r, fig.width = 6, fig.height = 4}
bind_rows(yeast = fdrcontrol_yeast, sims = fdrcontrol_sims, .id = "analysis") %>%
    dplyr::mutate(Method = reorder(Method, pExceeds)) %>%
    ggplot(aes(x = Method, y = pExceeds, color = Method, group = analysis)) +
    geom_line(aes(linetype = analysis), color = 'black', alpha = 1/4) +
    geom_point() + 
    scale_color_manual(values = col, guide = FALSE) +
    scale_x_discrete(NULL) + 
    scale_y_continuous("Percent of settings exceeding nominal FDR", labels = scales::percent) + 
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14)) +
    ggtitle("FDR control across simulations and yeast experiments") 
```

@@

```{r, fig.width = 6, fig.height = 4}
bind_rows(yeast = power_yeast, sims = power_sims, .id = "analysis") %>%
    dplyr::mutate(Method = reorder(Method, meanRank)) %>%
    ggplot(aes(x = Method, y = meanRank, color = Method, group = analysis)) +
    geom_line(aes(linetype = analysis), color = 'black', alpha = 1/4) +
    geom_point() + 
    scale_color_manual(values = col, guide = FALSE) +
    scale_x_discrete(NULL) + 
    scale_y_continuous("Mean TPR rank", breaks = 0:10) + 
    expand_limits(y = 1) +
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14)) +
    ggtitle("TPR rank across simulations and yeast experiments")

bind_rows(yeast = power_yeast, sims = power_sims, .id = "analysis") %>%
    dplyr::mutate(Method = reorder(Method, medianRank)) %>%
    ggplot(aes(x = Method, y = medianRank, color = Method, group = analysis)) +
    geom_line(aes(linetype = analysis), color = 'black', alpha = 1/4) +
    geom_point() + 
    scale_color_manual(values = col, guide = FALSE) +
    scale_x_discrete(NULL) + 
    scale_y_continuous("Median TPR rank", breaks = 0:10) + 
    expand_limits(y = 1) +
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14)) +
    ggtitle("TPR rank across simulations and yeast experiments") 
```

@@

```{r, fig.width = 6, fig.height = 6}
bind_rows(yeast = robustness_yeast, sims = robustness_sims,
          casestudies = robustness_data, .id = "analysis") %>%
    dplyr::mutate(Method = reorder(Method, sdGain)) %>%
    ggplot(aes(x = Method, y = sdGain, color = Method, group = analysis)) +
    geom_line(aes(linetype = analysis), color = 'black', alpha = 1/4) +
    geom_point() + 
    scale_color_manual(values = col, guide = FALSE) +
    scale_x_discrete(NULL) + 
    scale_y_continuous("SD of percent relative to BH/qvalue") + 
    expand_limits(y = 1) +
    facet_grid(key ~ ., scales = "free") + 
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14)) +
    ggtitle("Variation across simulations, yeast experiments,\ncase studies")

@@

We also include data sets where 0 rejections were made by `adapt-glm` as cases where
the method could not be applied.

```{r}
applicability_data <- datres %>%
    dplyr::mutate(nrej = ifelse(nrej == 0 & Method == "adapt-glm", NA, nrej)) %>%
    dplyr::group_by(casestudy, Method) %>%
    dplyr::summarize(meanNA = mean(is.na(nrej))) %>%
    dplyr::group_by(Method) %>%
    dplyr::summarize(meanNA = mean(meanNA)) %>%
    dplyr::ungroup()
dplyr::arrange(applicability_data, meanNA)
```
